---
title: "Figures"
author: "Lin Zhang"
output: 
  html_document: 
    keep_md: yes
---

Marmoset Paper figures
# Fig1
Fig1f heatmap of cell class specific markers in neonatal marmoest fovea
```{r}
Retina_cellclass_markers <- c("GNAT1","RHO","SAG", # Rod markers
                              "GNAT2","ARR3","GNGT2",# Cone markers
                              "ONECUT1","ONECUT2","LHX1","RET", # HC markers
                              "VSX2","VSX1","GRM6","CABP5","TMEM215","CHTF18", # BC markers
                              "SLC32A1","TFAP2A","TFAP2B","GAD1", # AC markers
                              "RBPMS","SLC17A6",# RGC markers
                              "APOE","SLC1A3","RLBP1","CRYAB" # MG markers
                              )
```

```{r}
P0_fovea <- readRDS("P0_Fovea.rds")
```

```{r}
DoHeatmap(P0_fovea, features = Retina_cellclass_markers)
```

# Fig2
## Fig2a dimplots of adult fovea atlas
### PR
```{r}
(Adult_F_PR <- readRDS("Adult_Fovea_PR.rds"))
```
```{r}
DimPlot(Adult_F_PR,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### HC
```{r}
(Adult_F_HC <- readRDS("Adult_Fovea_HC.rds"))
```
```{r}
DimPlot(Adult_F_HC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### BC
```{r}
(Adult_F_BC <- readRDS("Adult_Fovea_BC.rds"))
```
```{r}
DimPlot(Adult_F_BC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### AC
```{r}
(Adult_F_AC <- readRDS("Adult_Fovea_AC.rds"))
```
```{r}
DimPlot(Adult_F_AC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### RGC
```{r}
(Adult_F_RGC <- readRDS("Adult_Fovea_RGC.rds"))
```
```{r}
DimPlot(Adult_F_RGC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### Non neuronal cells
```{r}
(Adult_F_NN <- readRDS("Adult_Fovea_NN.rds"))
```
```{r}
DimPlot(Adult_F_NN,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

## Fig2b dimplots of adult periphery cell atlas

### PR
```{r}
(Adult_P_PR <- readRDS("Adult_Periphery_PR.rds"))
```
```{r}
DimPlot(Adult_P_PR,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### HC
```{r}
(Adult_P_HC <- readRDS("Adult_Periphery_HC.rds"))
```
```{r}
DimPlot(Adult_P_HC,reduction="umap",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### BC
```{r}
(Adult_P_BC <- readRDS("Adult_Periphery_BC.rds"))
```
```{r}
DimPlot(Adult_P_BC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### AC
```{r}
(Adult_P_AC <- readRDS("Adult_Periphery_AC.rds"))
```
```{r}
DimPlot(Adult_P_BC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### RGC
```{r}
(Adult_P_RGC <- readRDS("Adult_Periphery_RGC.rds"))
```
```{r}
DimPlot(Adult_P_RGC,reduction="umap",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### Non neuronal cells
```{r}
(Adult_P_NN <- readRDS("Adult_Periphery_NN.rds"))
```
```{r}
DimPlot(Adult_P_NN,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

## Fig2c dotplots of cell class- and type-specific markers

#### cell class specific markers
```{r}
PR_Panmarker <- c("ARR3","GNAT1")  
HC_Panmarker <- c("ONECUT1")   
BC_Panmarker <- c("VSX2")     
AC_Panmarker <- c("SLC32A1") 
RGC_Panmarker <- c("RBPMS")
MG_Panmarker <- c("VIM")
```

#### cell type specific markers
```{r}
# PR
P0_F_PR <- readRDS(file= "P0_Fovea_PR.rds")
MarkergeneHeatmap(data = P0_F_PR, output = P0_F_PR.markers, file = "P0_F_PR", ne = 3)

# HC
P0_F_HC <- readRDS(file= "P0_Fovea_HC.rds")
MarkergeneHeatmap(data = P0_F_HC, output = P0_F_HC.markers, file = "P0_F_HC", ne = 3)

# BC
P0_F_BC <- readRDS(file= "P0_Fovea_BC.rds")
MarkergeneHeatmap(data = P0_F_BC, output = P0_F_BC.markers, file = "P0_F_BC", ne = 3)

# AC
P0_F_AC <- readRDS(file= "P0_Fovea_AC.rds")
MarkergeneHeatmap(data = P0_F_AC, output = P0_F_AC.markers, file = "P0_F_AC", ne = 3)

# RGC
P0_F_RGC <- readRDS(file = "P0_Fovea_RGC.rds")
MarkergeneHeatmap(data = P0_F_RGC, output = P0_F_RGC.markers, file = "P0_F_RGC", ne = 3)

# Non-neuronal cells
P0_F_NN <- readRDS(file = "P0_Fovea_NN.rds")
MarkergeneHeatmap(data = P0_F_NN, output = P0_F_NN.markers, file = "P0_F_NN", ne = 3)
```

#### dotplot
```{r}
## load identified marker genes P0_F_marker_genes
load("P0_F_marker_genes.rda")

## merge p0 fovea objects
P0_F_Merged <- merge(P0_F_PR, y = c(P0_F_HC,P0_F_BC, P0_F_AC, P0_F_RGC, P0_F_NN), add.cell.ids=c("P0_F_PR","P0_F_HC", "P0_F_BC","P0_F_AC","P0_F_RGC", "P0_F_NN"),project="P0_F_Merged")

## only MG of non-neuronal cells 
P0_F_Merged <- subset(P0_F_Merged, idents = c("Pericyte","Microglia","Endothelial"), invert = TRUE)

DotPlot(P0_F_Merged, features = P0_F_marker_genes, cols = c("white", "#08306B"),dot.scale = 12, col.min = 0, col.max = 2)+coord_flip()+ RotatedAxis()
```

# Fig3
## Fig3a
Circos from inner layer to outer layer
### umap
```{r}
PrimateThree <- readRDS(file="PrimateThree_Integrated.rds")
```

```{r}
DimPlot(PrimateThree,reduction="umap",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### dendrogram
```{r}
PrimateThree <- BuildClusterTree(PrimateThree,assay = "integrated",slot = "scale.data")
```
```{r,fig.height=5}
data.tree <- Tool(object=PrimateThree_V2,slot='BuildClusterTree')
plot.phylo(data.tree)
```

### dotplot
```{r,fig.height=12,fig.width=15}
(p1 <- DotPlot(PrimateThree,features = unique(PerCellClassTop3Markers),cols=c("white","red"),dot.scale = 8,split.by = "species")+RotatedAxis())
```

### barplot
```{r}
cellclass <- c("Non-neural","RGC","AC&HC","BC","PR")
```

```{r}
prop_df$species <- as.factor(prop_df$species)
```
```{r}
color <- c(1:3)
```

```{r}
library(circlize)
```

```{r}
circos.clear()
circos.initialize(sectors = cellclass, xlim = XLIM)
circos.track(ylim = c(0, 1), panel.fun = function(x, y) {
    value = prop_df$Nor_prop[prop_df$cellclass==CELL_META$sector.index]
    circos.barplot(value, col =as.integer(prop_df$species[prop_df$cellclass==CELL_META$sector.index]))
})
#1: black, human 2:red,macaque 3:green marmoset
```

## Fig3b
```{r,fig.height=0.6,fig.width=4}
pheatmap(Corr_Mat,cluster_rows = FALSE,cluster_cols = TRUE)
```

# Fig4
## Fig4a dimplots of neonatal fovea atlas
### PR
```{r}
(P0_F_PR <- readRDS("P0_Fovea_PR.rds"))
```

```{r}
DimPlot(P0_F_PR,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### HC
```{r}
(P0_F_HC <- readRDS("P0_Fovea_HC.rds"))
```

```{r}
DimPlot(P0_F_HC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### BC
```{r}
(P0_F_BC <- readRDS("P0_Fovea_BC.rds"))
```

```{r}
DimPlot(P0_F_BC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### AC
```{r}
(P0_F_AC <- readRDS("P0_Fovea_AC.rds"))
```

```{r}
DimPlot(P0_F_AC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### RGC
```{r}
(P0_F_RGC <- readRDS("P0_Fovea_RGC.rds"))
```

```{r}
DimPlot(P0_F_RGC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### Non-neuronal
```{r}
(P0_F_NN <- readRDS("P0_Fovea_NN.rds"))
```

```{r}
DimPlot(P0_F_NN,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

## Fig4b dimplots of neonatal periphery atlas
### PR
```{r}
(P0_P_PR <- readRDS("P0_Periphery_PR.rds"))
```

```{r}
DimPlot(P0_P_PR,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### HC
```{r}
(P0_P_HC <- readRDS("P0_Periphery_HC.rds"))
```

```{r}
DimPlot(P0_P_HC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### BC
```{r}
(P0_P_BC <- readRDS("P0_Periphery_BC.rds"))
```

```{r}
DimPlot(P0_P_BC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### AC
```{r}
(P0_P_AC <- readRDS("P0_Periphery_AC.rds"))
```

```{r}
DimPlot(P0_P_AC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### RGC
```{r}
(P0_P_RGC <- readRDS("P0_Periphery_RGC.rds"))
```

```{r}
DimPlot(P0_P_RGC,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

### Non neuronal cells
```{r}
(P0_P_NN <- readRDS("P0_Periphery_NN.rds"))
```

```{r}
DimPlot(P0_P_NN,reduction="UMAP",label=TRUE,label.size=3,repel=TRUE)+NoLegend()
```

## Fig4c dotplots of cell class- and type-specific markers
#### cell class specific markers
```{r}
PR_Panmarker <- c("ARR3","GNAT1")  
HC_Panmarker <- c("ONECUT1")   
BC_Panmarker <- c("VSX2")     
AC_Panmarker <- c("SLC32A1") 
RGC_Panmarker <- c("RBPMS")
MG_Panmarker <- c("VIM")
```

#### cell type specific markers
```{r}
# PR
Adult_F_PR <- readRDS(file= "Adult_Fovea_PR.rds")
MarkergeneHeatmap(data = Adult_F_PR, output = Adult_F_PR.markers, file = "Adult_F_PR", ne = 3)

# HC
Adult_F_HC <- readRDS(file= "Adult_Fovea_HC.rds")
MarkergeneHeatmap(data = Adult_F_HC, output = Adult_F_HC.markers, file = "Adult_F_HC", ne = 3)

# BC
Adult_F_BC <- readRDS(file= "Adult_Fovea_BC.rds")
MarkergeneHeatmap(data = Adult_F_BC, output = Adult_F_BC.markers, file = "Adult_F_BC", ne = 3)

# AC
Adult_F_AC <- readRDS(file= "Adult_Fovea_AC.rds")
MarkergeneHeatmap(data = Adult_F_AC, output = Adult_F_AC.markers, file = "Adult_F_AC", ne = 3)

# RGC
Adult_F_RGC <- readRDS(file = "Adult_Fovea_RGC.rds")
MarkergeneHeatmap(data = Adult_F_RGC, output = Adult_F_RGC.markers, file = "Adult_F_RGC", ne = 3)

# Non-neuronal cells
Adult_F_NN <- readRDS(file = "Adult_Fovea_NN.rds")
MarkergeneHeatmap(data = Adult_F_NN, output = Adult_F_NN.markers, file = "Adult_F_NN", ne = 3)
```

#### dotplot
```{r}
## load identified marker genes P0_F_marker_genes
load("Adult_F_marker_genes.rda")

## merge Adult fovea objects
Adult_F_Merged <- merge(Adult_F_PR, y = c(Adult_F_HC,Adult_F_BC, Adult_F_AC, Adult_F_RGC, Adult_F_NN), add.cell.ids=c("Adult_F_PR","Adult_F_HC", "Adult_F_BC","Adult_F_AC","Adult_F_RGC", "Adult_F_NN"),project="Adult_F_Merged")

## only MG of non-neuronal cells 
Adult_F_Merged <- subset(Adult_F_Merged, idents = c("Pericyte","Microglia","Endothelial"), invert = TRUE)

DotPlot(Adult_F_Merged, features = Adult_F_marker_genes, cols = c("white", "#08306B"),dot.scale = 12, col.min = 0, col.max = 2)+coord_flip()+ RotatedAxis()
```

# Fig5
## Fig5a P0 Fovea to P0 Periphery
```{r}
library(circlize)
```

```{r,fig.height=8,fig.width=8}
circos.clear()
circos.par(start.degree = 90)

chordDiagram(C,order = c(Fovea_order,rev(Per_order)),annotationTrackHeight = c(0.05,0.05),annotationTrack = "grid") 

circos.track(track.index = 1, panel.fun = function(x, y) {
    circos.text(CELL_META$xcenter, CELL_META$ylim[1],
        facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5),cex = 0.4)
}, bg.border = NA)
```

## Fig5b Adult Fovea to Adult Periphery
```{r}
library(circlize)
```

```{r,fig.height=8,fig.width=8}
circos.clear()
circos.par(start.degree = 90)

chordDiagram(C,order = c(Fovea_order,rev(Per_order)),annotationTrackHeight = c(0.05,0.05),annotationTrack = "grid") 

circos.track(track.index = 1, panel.fun = function(x, y) {
    circos.text(CELL_META$xcenter, CELL_META$ylim[1],
        facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5),cex = 0.4)
}, bg.border = NA)
```

## Fig5c Neonatal Fovea to Adult Fovea
```{r}
library(circlize)
```

```{r,fig.height=8,fig.width=8}
circos.clear()
circos.par(start.degree = 90)

chordDiagram(C,order = c(Adult_order,rev(P0_order)),annotationTrackHeight = c(0.05,0.05),annotationTrack = "grid") 

circos.track(track.index = 1, panel.fun = function(x, y) {
    circos.text(CELL_META$xcenter, CELL_META$ylim[1],
        facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5),cex = 0.4)
}, bg.border = NA)
```

## Fig5d Neonatal Periphery to Adult Periphery
```{r}
library(circlize)
```

```{r,fig.height=8,fig.width=8}
circos.clear()
circos.par(start.degree = 90)

chordDiagram(C,order = c(Adult_order,rev(P0_order)),annotationTrackHeight = c(0.05,0.05),annotationTrack = "grid") 

circos.track(track.index = 1, panel.fun = function(x, y) {
    circos.text(CELL_META$xcenter, CELL_META$ylim[1],
        facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5),cex = 0.4)
}, bg.border = NA)
```

# Fig6
## Fig6a
```{r}
(P0_F_P <- readRDS(file="P0_Fovea_and_Periphery.rds"))
```
```{r}
FeaturePlot(P0_F_P,features=c("NES"),reduction = "UMAP",label = TRUE,repel = TRUE,max.cutoff = 25,label.size = 5)
```

## Fig6b
```{r}
df$EMD_scale <- scale(df$EMD)
```

```{r}
p <- ggplot(df,aes(x=reorder(cell,EMD_scale),y=EMD_scale))+geom_bar(stat="identity",size=0.8,color="black",alpha=0.3)+coord_flip()+
  geom_col(aes(fill = cellclass)) 

image = p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), legend.title = element_blank(),axis.line = element_line(colour = "black")) 

image
```

## Fig6c
```{r}
(Adult_F_P <- readRDS(file="Adult_Fovea_and_Periphery.rds"))
```
```{r}
FeaturePlot(Adult_F_P,features=c("NES"),reduction = "UMAP",label = TRUE,repel = TRUE,max.cutoff = 25,label.size = 5)
```

## Fig6d
```{r}
df$EMD_scale <- scale(df$EMD)
```

```{r}
p <- ggplot(df,aes(x=reorder(cell,EMD_scale),y=EMD_scale))+geom_bar(stat="identity",size=0.8,color="black",alpha=0.3)+coord_flip()+
  geom_col(aes(fill = cellclass)) 

image = p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), legend.title = element_blank(),axis.line = element_line(colour = "black")) 

image
```

## Fig6e
```{r}
(Fovea_P0_Adult <- readRDS(file="Fovea_P0_and_Adult.rds"))
```
```{r}
FeaturePlot(Fovea_P0_Adult,features=c("NES"),reduction = "UMAP",label = TRUE,repel = TRUE,max.cutoff = 25,label.size = 5)
```

## Fig6f
```{r}
df$EMD_scale <- scale(df$EMD)
```

```{r}
p <- ggplot(df,aes(x=reorder(cell,EMD_scale),y=EMD_scale))+geom_bar(stat="identity",size=0.8,color="black",alpha=0.3)+coord_flip()+
  geom_col(aes(fill = cellclass)) 

image = p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), legend.title = element_blank(),axis.line = element_line(colour = "black")) 

image
```

## Fig6g
```{r}
(Periphery_P0_Adult <- readRDS(file="Periphery_P0_and_Adult.rds"))
```
```{r}
FeaturePlot(Periphery_P0_Adult,features=c("NES"),reduction = "UMAP",label = TRUE,repel = TRUE,max.cutoff = 25,label.size = 5)
```

## Fig6h
```{r}
df$EMD_scale <- scale(df$EMD)
```

```{r}
p <- ggplot(df,aes(x=reorder(cell,EMD_scale),y=EMD_scale))+geom_bar(stat="identity",size=0.8,color="black",alpha=0.3)+coord_flip()+
  geom_col(aes(fill = cellclass)) 

image = p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), legend.title = element_blank(),axis.line = element_line(colour = "black")) 

image
```

# Fig7
Fig7a
```{r}
(Marmoset_MLCone <- readRDS(file="Marmoset_MLCone.rds"))
```

```{r}
DimPlot(Marmoset_MLCone,reduction = "umap",pt.size = 0.5)+theme(aspect.ratio = 1)
```

# Fig8
Fig8c
```{r}
ligand_target_matrix = readRDS("ligand_target_matrix.rds")
lr_network = readRDS("lr_network.rds")
weighted_networks = readRDS("weighted_networks.rds")
```

```{r}
p_ligand_target_network = vis_ligand_target %>% make_heatmap_ggplot("Prioritized ligands enriched in foveal MG","Neonatal Fovea enriched DEGs of M/L-Cones", color = "darkblue",legend_position = "bottom", x_axis_position = "top",legend_title = "Scaled Regulatory potential") + scale_fill_gradient2( low = muted("green"),mid = "white",high = muted("darkred"),na.value = "white smoke",breaks = c(-2,0,2)) + theme(axis.text.x = element_text(face = "italic"))

p_ligand_target_network
```

